Combination of field and experimental data together with computational models reveal cognitive mechanisms behind cleaner fish behaviour

### 2021-12-21 manuscript_1.0/manuscript_1.0.tex
Copies repo from Clean.ActCrit this will be released with manuscript forthcoming

---
title: Combination of field and experimental data together with computational models 
  reveal cognitive mechanisms behind cleaner fish behaviour
#Cognitive and ecological aspects determine cleaner performance
author:
  - name: Andrés E. Quiñones 
    affiliation: "1"
  - name: Zegni Triki
    affiliation: "2"
  - name: Redouan Bshary
    affiliation: "1"
address:
  - code: "1"
    address: Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland
  - code: "2"
    address: Department of Zoology, Stockholm University, Stockholm, Sweden

corresp_author_name:  "AE. Quiñones"
corresp_author_email: "andreseqp@gmail.com"

subject:
  - "Behavioural ecology"
  - "Cognitive ecology"
  - "Animal behaviour"

keywords:
  - learning   
  - cleaners
  - behaviour
  - bayesian statisitics
  - mutualism

abstract: |
  The same behaviour can be achieved by different mechanistic underpinnings. But the environmental conditions under which the behaviour is triggered might differ between these potential mechanisms. Thus, by analyzing the varitation of expression of a certain behavior across environmental gradients, it is possible to infer which mechanism is behind the behaviour of interest. To do this it is necessary to have clearcut predictions for the different candidate mechanisms. In here, we use experimental data on the behaviour of the cleaner fish *Labroides* *dimidiatus*, as well as enviromental variables from the location of fish, together with a computational model of the learning mechanisms. Using bayesian statistics, we fit key parameters of the computational model that determine the specific learning mechanism. We find that a model where cleaners estimate the long-term consequences of their actions, and use the estimate in their desicion making, explains better under which conditions the fish develop a preference for an ephemeral food source in the context of a biological market. We argue that the combination of computational models with data is powerful tool to infer the mechanistic underpinning behind behaviours of interest.

## Remove this if not required
ethics: |
  The Animal Ethics Committee of the Queensland government (DAFF)
  approved the project (CA 2016/05/970 and CA 2017/05/1063).

data_accessibility: |
  Please provide details on the data availability.

author_contributions: |
  AQ and RB designed the study. ZT collected the data. AQ developed the model and 
  fitted the model parameters to the data. AQ, RB and ZT wrote the manuscript.

## Remove this if not required
conflict_of_interest: |
  The authors declate no conflict of interests.

## Remove this if not required
funding: |
  This work was supported by the Swiss National Science Foundation (grants
  number: 31003A_153067/1 and 310030B_173334/1 to R.B.).

## Remove this if not required
acknowledgements: |
  ZT and RB kindly thank the staff of Lizard Island Research Station, Dominique
  Roche, Sharon Wismer, Olivia Rey,Sandra Ann Binning, Elena Levorato and William McNeely for their
  field support. AQ thanks Florian Hartig for statistical advice.

bibliography: Cleanerlearning.bib

## change to true to add optional line numbering
lineno: true

header-includes: |
  ```{=latex}
  \usepackage{float}
  \usepackage{booktabs}
  \newcommand{\beginsupplement}{ \setcounter{table}{0}     \renewcommand{\thetable}{S\arabic{table}}\setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}
  ```
site: bookdown::bookdown_site
output: 
  bookdown::pdf_book:
    base_format: rticles::rsos_article
---

## Introduction

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE,warning = FALSE,message=FALSE)
library(here)
```

Cooperation among individuals of the same or different species is
ubiquitous in nature. Among individuals of different species it
underlays the biological mechanism for fundamental ecological
interactions such as pollination, nitrogen fixation or parasite removal
[@begon_Ecology_2006]. When partners in these mutualistic interactions
are able to choose their interaction partner, their investment and the
return they obtain can be understood as a biological market
[@hammerstein_Biological_2016; @leimar_Cooperation_2010;
@noe_Biological_1994a;@noe_Biological_1995a]. Thus, we expect supply
and demand of the good and services being exchanged to determine the
action of individuals, at least in an evolutionary time-scale. However,
supply and demand conditions of a biological market, which typically
depend on the abundance of the species involved, certainly vary in time
and space. How do partners adjust their strategy in the face of this
variation? In animals an obvious general candidate mechanism for the
strategic adjustment is the cognitive machinery. However, it is not
clear which specific cognitive mechanisms allow individual to adjust
their behaviour to the varying market conditions.

One example of strategic adjustment in a biological market is the marine
cleaning mutualism involving the cleaner fish *Labroides dimidiatus* and
'client' fish. Clients offer themselves as food patches for cleaners to
remove ectoparasites, which gives the cleaners calories and the clients
improved health [@waldie_LongTerm_2011;
@ros_Does_2011;@triki_Effects_2016]. As cleaners have about 2000
interactions per day [@grutter_Relationship_1995], they regularly face
the situation that two clients seek cleaning simultaneously. In such
conditions, it is important to distinguish between 'resident' and
'visitor' client species [@bshary_Cleaner_2001a]. The residents have
small home ranges and hence access to only the local cleaner, while the
visitors have large home ranges and hence access to two or more
cleaners. Thus, visitors have the option to switch to another cleaner if
being made to wait, while residents must wait for inspection. Indeed,
visitors have been observed to use their partner choice options in that
way [@bshary_Choosy_2002], which may explain why cleaners give
visitors service priority in a field study in the Red Sea
[@bshary_Cleaner_2001a]. More generally, cleaners apparently prefer to
interact with visitors [@triki_Marine_2021]. Such a preference may be
linked to a client's value as a food patch; visitors are on average
larger than residents [ @bshary_Cleaner_2001a], and size is positively
correlated with parasite load [@grutter_Relationship_1995a]. However, 
early laboratory
experiments showed that leaving causes service priority independent of
client value [@bshary_Asymmetric_2002]. The experiments consisted of a
simultaneous two-choice task, where cleaners had to decide which of two
Plexiglas plates, which offered equal amounts of food, to approach first
and eat mashed prawn of it. One plate was made to behave like a
resident, i.e. it would remain until the cleaner had eaten off it. The
other plate was made to behave like a visitor, i.e. it would be removed
if not inspected first. Cleaners learned to prefer the visitor plate and
hence obtained the double amount of food [@bshary_Asymmetric_2002].
Adult cleaner fish outperform various primates as well as rats and
pigeons in this original version of the biological market task
[@salwiczek_Adult_2012; @zentall_Early_2017]; only African grey
parrots solve this task as well [@pepperberg_Can_2014].

Recent research has refined our view on partner choice and priority of
access in the *L. dimidiatus* cleaning mutualism, and as a consequence
performance in the biological market task. First, details of the task
presentation may matter; i.e. what relevant information is encoded in
the design of the two plates (colour versus size;
[@wismer_Cuebased_2019]), or whether the information is in the colour of
the food [@pretot_Comparative_2021; @pretot_Comparing_2016], and
whether the food is visible or hidden/released only when the subjects
made choices on a touch screen [@pretot_Factors_2016]. Second, it
appears that cleaner and client densities in nature affect how cleaners
perform in the original market task: when densities are low, cleaners
are more likely to fail to complete the task [@triki_Biological_2019;
@triki_Decrease_2018; @wismer_Variation_2014]. The apparent explanation
for this failure is that visitor clients are less likely to exert their
choice options (and hence wait for inspection if made to wait) when
cleaner fish densities are low [@triki_Brain_2020]. This change in
behaviour may be due to the fact that when cleaner densities are low,
the cost of switching in terms of distance to swim and time invested
increases. If visitors are not choosy, cleaners should not give them
priority anymore. Interestingly, while cleaners from high-density areas
have larger forebrains harbouring more cells than cleaners from
low-density areas [@triki_Brain_2020], on a local scale those
individuals that perform in the market task in a way that corresponds to
the locally optimal strategy (solving/failing the task when living in
high/low density) have relatively larger forebrains [@triki_Brain_2020].
Taken together, empirical results clearly show major intra- and
inter-specific variation in performance in the market task, warranting
investigation of the underlying cognitive machinery, from perception to
decision-making [@shettleworth_Cognition_2009].

One approach to identify candidate mechanisms underlying performance in
a cognitive task is to develop proximate learning models and to evaluate
their predictions in light of the behavioural data (e.g.
[@dubois_Model_2021]). By varying the cognitive tool kit, one can assess
which minimal kit is necessary to solve the task at hand. This approach
has been used to analyse the market task [@prat_Role_2021;
@quinones_Reinforcement_2019]. Learning models showed that basic
reinforcement learning does not suffice to solve the task, this is
particularly so when models assume the more complex natural situation in
which cleaners not only face resident-visitor pairs but also
visitor-visitor and resident-resident pairs, as well only a resident or
only a visitor. To be able to give visitors priority over residents,
cleaners first need to be able to assess a client's value separately for
the three possible forms of encounter (alone, paired with a fish with
the same strategic option, paired with a fish with the alternative
strategic option) [@quinones_Reinforcement_2019]. In addition, cleaners
need to account for the future consequences of current decisions. In the
model, this could be achieved in two non-mutually exclusive ways
[@quinones_Reinforcement_2019]: through low temporal discounting of
future effects and/or through perceiving a visitor client leaving as
psychological punishment (i.e. as a negative reinforcer). Given that a
negative reinforcer is part of the ubiquitous learning mechanism termed
operant conditioning [@thorndike_Animal_1898; @skinner_Behavior_1938], 
this explanation can be
considered the simpler one. Furthermore, accounting for the future
effects of an action is considered an advanced cognitive ability
[@suddendorf_Evolution_2007]. Here, we fitted our model to the empirical
data on fish densities and cleaner fish performance in the market task
with the aim to determine the likely mechanism used by cleaners to
incorporate future consequences of current decisions.

## Methods

### Overview

We used the data from [@triki_Biological_2019] to fit the parameters of
the model analysed in [@quinones_Reinforcement_2019]. For the bayesian
fitting procedure we used Markov Chain Monte Carlo (MCMC) with a
Metropolis sampler to estimate the posterior distribution of the model
parameters. The data consist a of the number of times that a set of
cleaners chose to feed of a visitor client plate in the context of the
market task. This set of cleaners were collected in different locations
and in each one of those locations environmental variables were
measured. These environmental variables recorded from each location are
used as input for the computational model. Given these inputs, the
computational model produces a time series of choices in the market
task. In the model the preference for the two options changes as the
algorithm learns based on the reward that the two options provide. Thus,
we run the model for 10000 steps and use the last 3000 steps as the
preference developed in the model for visitors given the environmental
variables. We use this preference as the predicted probability in a
binomial distribution to calculate the likelihood that the number of
visitor choices obtained from each cleaner is generated by the model. We
sum over all the cleaners the log-likehoods and use the total for the
acceptance of new parameters in the MCMC.

### The model

The model consist of a set individual-based simulations where
individuals face a series of choices between two options. Individuals in
the model represent cleaners, and the options they face simulate the
natural conditions of a cleaning market. Thus, individuals in the model
choose to prioritize different clients. These clients are classified as
resident and visitors. Residents are those clients that stay in the
cleaning queue when they are not given priority; while visitors leave
the queue (with a certain probability) after they are not given
priority. Individuals obtain a fixed reward from cleaning a client
regardless of the type. Every time individuals face and make a choice
they update the probability of making that same choice. The update is
based on the difference between the expectation of reward and the actual
reward the individual obtains (prediction error)
[@sutton_Reinforcement_2018; @rescorla_Theory_1972]; and it is carried
in the direction that would lead to more reward being obtained, given
the new information. Further details of the model implementation can be
found in the paper [@quinones_Reinforcement_2019]. The model analysis
points to the relevance of two parameters. First, $\gamma$ measures how
much individuals include future rewards in their decision updates. If
$\gamma=0$, individuals only use the immediate reward obtained from a
cleaning interaction. As $\gamma$ increases individuals include more the
reward obtained from the subsequent choices. That amounts to calculating
and using for decision making the future expected rewards in action.
Second, $\eta$ measures how much individuals include in their reward the
fleeing behaviour of visitor clients as a negative component. Both of
these parameters allow individuals to use in their estimates the future
effects of their choices. The model analysis by Quiñones *et al*
[@quinones_Reinforcement_2019] shows that without either of them, learning
individuals cannot develop a preference for the visitor client. The
analysis contrast the effect of these two parameters against variation
in the relative abundance of the two client types. The analysis shows
that under very high client (relative) abundance the presence of future
rewards ($\gamma>0$) no longer favours the preference for the ephemeral
clients, this effect does not occur with negative reward mediated by
client behaviour ($\eta$). Thus, the analysis suggests that future
reward as an explanation for cleaner ability to prefer visitors is
supported by Triki *et al* [@triki_Biological_2019] data. However, a combination of
future and negative rewards, could still be a better explanation for the
data. In the current analysis, we will use the data from
[@triki_Biological_2019] to fit some of the model parameters, particularly
relevant we will fit $\gamma$ and $\eta$ to find out whether each or a
combination of these effects is a better explanation for the pattern
seen in the data. For other parameter values used for the model
simulations and the MCMC see Table \ref{tab:param}

### Data explanation

*Insert here brief description of the data and data collection*

In order to capture with the model the relationship between the
characteristics of the market (environment) and the cleaner's
performance we need to translate the absolute abundances of the data to
to  a measure of relative abundances that captures client visitation patterns.
This is because in the model, relative abundances
of clients define not only the probability of residents and visitors,
but also how often the cleaning station is empty (e.g. there are no
clients to be cleaned). The more cleaners there are, the more options
are available for the clients and hence the higher is the probability of
empty cleaning stations. How often the cleaning stations are empty in
turns influences the value that cleaners can expect from their choices.
The emptier the cleaning stations are, the more beneficial it is to
prioritize a client that will not stay. However, the frequency of empty
spots in the cleaning station does not only depend on cleaner abundances 
but also on the frequency with which clients visit the
station. The client visitation frequency will likely vary among
different clients depending on parasite abundances and loads. These are
variables for which we do not have field estimates. Thus, we computed a
measure of relative cleaner abundance for each site relative to absolute
abundances and multiplied by a scaling constant that changes the range
of the variable. This scaling constant is meant to capture variation in the market 
conditions driven partly by cleaner abundance. We fit the value of the 
scaling constant as part of the
statistical inference. As for the client abundances for residents and
visitors we computed a relative measure with respect to the total client
abundance, and weighted that by the rescaled relative absence of
cleaners (1-relative cleaner abundance). Thus, all three measures of
relative abundance sum up to 1, and can be used in the model as a proxy
for the probability of having different options in the cleaning station.

## Results

```{r post, fig.cap="Posterior distributions for parameter values $\\gamma$, $\\eta$ and the scaling constant. We show the kernel density estimates, below the mode (black point) and the 65\\% (light blue) and 95\\% (grey)  highest posterior density interval for the three paramteres. On the left (panels a, c and c) posterior distributions for a model with   all three parameters; while on the right (b and d) a model without negative reward. Panel f shows a measure of fit for both models, namely the distribution of pseudo-$R^2$ obtained from sampling the posterior distribution of parameter values.",out.width='95%',fig.height=9}
require(data.table)
library(coda)
library(ggplot2)
library(cowplot)
library(ggdist)
source(here("../R_files/posPlots.R"))
source(here("loadData.R"))
library("jsonlite")

# Name of the folders containing the MCMC runs
scen1<-"MCMCclean_gam_nRew_sca_"
scen2<-"MCMCclean_gam_sca_"
scen3<-"MCMCclean_Nrew_sca_"

# NLables names for the three scenarios
labels.Scen<-c("both","gam","Nrew")
labelsPlot.Scen<-c("Full model","Future reward","Negative reward")

## Load files --------------------------------------------------------------

# Defaults for MCMC analysis
burn.in<-1000
thinning<-100

MCMCdata1<-loadMCMCrun(scen1,thinning = thinning,burn.in = burn.in)
MCMCdata2<-loadMCMCrun(scen2,thinning = thinning,burn.in = burn.in)
MCMCdata3<-loadMCMCrun(scen3,thinning = thinning,burn.in = burn.in)

# Create list with the data, convenient to use CODA for diagnostics
mcmcList.1<-mcmc.list(do.call(list,lapply(unique(MCMCdata1$seed), function(repli){
  rundata<-MCMCdata1[seed==repli]
  mcmcRun<-mcmc(rundata[,.(gamma,negReward,scaleConst)],thin = thinning)#
  return(mcmcRun)
})))

mcmcList.2<-mcmc.list(do.call(list,lapply(unique(MCMCdata2$seed), function(repli){
  rundata<-MCMCdata2[seed==repli]
  mcmcRun<-mcmc(rundata[,.(gamma,scaleConst)],thin = thinning)#
  return(mcmcRun)
})))

mcmcList.2<-mcmc.list(do.call(list,lapply(unique(MCMCdata2$seed), function(repli){
  rundata<-MCMCdata2[seed==repli]
  mcmcRun<-mcmc(rundata[,.(negReward,scaleConst)],thin = thinning)#
  return(mcmcRun)
})))

mcmcList.3<-mcmc.list(do.call(list,lapply(unique(MCMCdata3$seed), function(repli){
  rundata<-MCMCdata3[seed==repli]
  mcmcRun<-mcmc(rundata[,.(gamma,negReward,scaleConst)],thin = 100)#
  return(mcmcRun)
})))

## Density plots with gg -------------------------------------------------------

## Define the boundaries for credible intervals using High Density Interval

intervals<-c(0.66,0.95)

cuts.1<-lapply(MCMCdata1[,.(gamma,negReward,scaleConst)],
             function(x){
               c(mode_hdci(x,.width = intervals[c(2,1)])$ymin,
                 mode_hdci(x,.width = intervals)$ymax)  
})

cuts.2<-lapply(MCMCdata2[,.(gamma,scaleConst)],
               function(x){
                 c(mode_hdi(x,.width = intervals[c(2,1)])$ymin,
                   mode_hdi(x,.width = intervals)$ymax)  
               })

cuts.3<-lapply(MCMCdata3[,.(negReward,scaleConst)],
               function(x){
                 c(mode_hdi(x,.width = intervals[c(2,1)])$ymin,
                   mode_hdi(x,.width = intervals)$ymax)  
               })

>>>>

%% Author_tex.tex
%% V1.0
%% 2012/13/12
%% developed by Techset
%%
%% This file describes the coding for rsproca.cls

\documentclass[]{rsos}%%%%where rsos is the template name

\usepackage{lineno}
\linenumbers

\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}

% For Pandoc highlighting needs

% Pandoc citation processing

\usepackage{float}
\usepackage{booktabs}
\newcommand{\beginsupplement}{ \setcounter{table}{0}     \renewcommand{\thetable}{S\arabic{table}}\setcounter{figure}{0} \renewcommand{\thefigure}{S\arabic{figure}}}

%%%% *** Do not adjust lengths that control margins, column widths, etc. ***

%%%%%%%%%%% Defining Enunciations  %%%%%%%%%%%
\newtheorem{theorem}{\bf Theorem}[section]
\newtheorem{condition}{\bf Condition}[section]
\newtheorem{corollary}{\bf Corollary}[section]
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

%%%% Article title to be placed here
\title{Combination of field and experimental data together with computational models reveal cognitive mechanisms behind cleaner fish behaviour}

\author{
Andrés E. Quiñones$^{1}$,
Zegni Triki$^{2}$,
Redouan Bshary$^{1}$}

\address{
  $^{1}$Institute of Biology, University of Neuchâtel, Neuchâtel, Switzerland\\
  $^{2}$Department of Zoology, Stockholm University, Stockholm, Sweden}
%%%% Subject entries to be placed here %%%%
\subject{
Behavioural ecology,
Cognitive ecology,
Animal behaviour}

%%%% Keyword entries to be placed here %%%%
\keywords{
learning,
cleaners,
behaviour,
bayesian statisitics,
mutualism}

%%%% Insert corresponding author and its email address}
\corres{
  AE. Quiñones\\
  e-mail: \href{mailto:andreseqp@gmail.com}{\nolinkurl{andreseqp@gmail.com}}
}

%%%% Abstract text to be placed here %%%%%%%%%%%%
\begin{abstract}
The same behaviour can be achieved by different mechanistic underpinnings. But the environmental conditions under which the behaviour is triggered might differ between these potential mechanisms. Thus, by analyzing the varitation of expression of a certain behavior across environmental gradients, it is possible to infer which mechanism is behind the behaviour of interest. To do this it is necessary to have clearcut predictions for the different candidate mechanisms. In here, we use experimental data on the behaviour of the cleaner fish \emph{Labroides} \emph{dimidiatus}, as well as enviromental variables from the location of fish, together with a computational model of the learning mechanisms. Using bayesian statistics, we fit key parameters of the computational model that determine the specific learning mechanism. We find that a model where cleaners estimate the long-term consequences of their actions, and use the estimate in their desicion making, explains better under which conditions the fish develop a preference for an ephemeral food source in the context of a biological market. We argue that the combination of computational models with data is powerful tool to infer the mechanistic underpinning behind behaviours of interest.
\end{abstract}
%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Some pieces required from the pandoc template
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\providecommand{\EndFirstPage}{%
}

\maketitle

\hypertarget{introduction}{%
\subsection{Introduction}\label{introduction}}

Cooperation among individuals of the same or different species is
ubiquitous in nature. Among individuals of different species it
underlays the biological mechanism for fundamental ecological
interactions such as pollination, nitrogen fixation or parasite removal
\citep{begon_Ecology_2006}. When partners in these mutualistic interactions
are able to choose their interaction partner, their investment and the
return they obtain can be understood as a biological market
\citep{hammerstein_Biological_2016, leimar_Cooperation_2010, noe_Biological_1994a, noe_Biological_1995a}. Thus, we expect supply
and demand of the good and services being exchanged to determine the
action of individuals, at least in an evolutionary time-scale. However,
supply and demand conditions of a biological market, which typically
depend on the abundance of the species involved, certainly vary in time
and space. How do partners adjust their strategy in the face of this
variation? In animals an obvious general candidate mechanism for the
strategic adjustment is the cognitive machinery. However, it is not
clear which specific cognitive mechanisms allow individual to adjust
their behaviour to the varying market conditions.

One example of strategic adjustment in a biological market is the marine
cleaning mutualism involving the cleaner fish \emph{Labroides dimidiatus} and
`client' fish. Clients offer themselves as food patches for cleaners to
remove ectoparasites, which gives the cleaners calories and the clients
improved health \citep{waldie_LongTerm_2011, ros_Does_2011, triki_Effects_2016}. As cleaners have about 2000
interactions per day \citep{grutter_Relationship_1995}, they regularly face
the situation that two clients seek cleaning simultaneously. In such
conditions, it is important to distinguish between `resident' and
`visitor' client species \citep{bshary_Cleaner_2001a}. The residents have
small home ranges and hence access to only the local cleaner, while the
visitors have large home ranges and hence access to two or more
cleaners. Thus, visitors have the option to switch to another cleaner if
being made to wait, while residents must wait for inspection. Indeed,
visitors have been observed to use their partner choice options in that
way \citep{bshary_Choosy_2002}, which may explain why cleaners give
visitors service priority in a field study in the Red Sea
\citep{bshary_Cleaner_2001a}. More generally, cleaners apparently prefer to
interact with visitors \citep{triki_Marine_2021}. Such a preference may be
linked to a client's value as a food patch; visitors are on average
larger than residents \citep{bshary_Cleaner_2001a}, and size is positively
correlated with parasite load \citep{grutter_Relationship_1995a}. However,
early laboratory
experiments showed that leaving causes service priority independent of
client value \citep{bshary_Asymmetric_2002}. The experiments consisted of a
simultaneous two-choice task, where cleaners had to decide which of two
Plexiglas plates, which offered equal amounts of food, to approach first
and eat mashed prawn of it. One plate was made to behave like a
resident, i.e.~it would remain until the cleaner had eaten off it. The
other plate was made to behave like a visitor, i.e.~it would be removed
if not inspected first. Cleaners learned to prefer the visitor plate and
hence obtained the double amount of food \citep{bshary_Asymmetric_2002}.
Adult cleaner fish outperform various primates as well as rats and
pigeons in this original version of the biological market task
\citep{salwiczek_Adult_2012, zentall_Early_2017}; only African grey
parrots solve this task as well \citep{pepperberg_Can_2014}.

Recent research has refined our view on partner choice and priority of
access in the \emph{L. dimidiatus} cleaning mutualism, and as a consequence
performance in the biological market task. First, details of the task
presentation may matter; i.e.~what relevant information is encoded in
the design of the two plates (colour versus size;
\citep{wismer_Cuebased_2019}), or whether the information is in the colour of
the food \citep{pretot_Comparative_2021, pretot_Comparing_2016}, and
whether the food is visible or hidden/released only when the subjects
made choices on a touch screen \citep{pretot_Factors_2016}. Second, it
appears that cleaner and client densities in nature affect how cleaners
perform in the original market task: when densities are low, cleaners
are more likely to fail to complete the task \citep{triki_Biological_2019, triki_Decrease_2018, wismer_Variation_2014}. The apparent explanation
for this failure is that visitor clients are less likely to exert their
choice options (and hence wait for inspection if made to wait) when
cleaner fish densities are low \citep{triki_Brain_2020}. This change in
behaviour may be due to the fact that when cleaner densities are low,
the cost of switching in terms of distance to swim and time invested
increases. If visitors are not choosy, cleaners should not give them
priority anymore. Interestingly, while cleaners from high-density areas
have larger forebrains harbouring more cells than cleaners from
low-density areas \citep{triki_Brain_2020}, on a local scale those
individuals that perform in the market task in a way that corresponds to
the locally optimal strategy (solving/failing the task when living in
high/low density) have relatively larger forebrains \citep{triki_Brain_2020}.
Taken together, empirical results clearly show major intra- and
inter-specific variation in performance in the market task, warranting
investigation of the underlying cognitive machinery, from perception to
decision-making \citep{shettleworth_Cognition_2009}.

One approach to identify candidate mechanisms underlying performance in
a cognitive task is to develop proximate learning models and to evaluate
their predictions in light of the behavioural data (e.g.
\citep{dubois_Model_2021}). By varying the cognitive tool kit, one can assess
which minimal kit is necessary to solve the task at hand. This approach
has been used to analyse the market task \citep{prat_Role_2021, quinones_Reinforcement_2019}. Learning models showed that basic
reinforcement learning does not suffice to solve the task, this is
particularly so when models assume the more complex natural situation in
which cleaners not only face resident-visitor pairs but also
visitor-visitor and resident-resident pairs, as well only a resident or
only a visitor. To be able to give visitors priority over residents,
cleaners first need to be able to assess a client's value separately for
the three possible forms of encounter (alone, paired with a fish with
the same strategic option, paired with a fish with the alternative
strategic option) \citep{quinones_Reinforcement_2019}. In addition, cleaners
need to account for the future consequences of current decisions. In the
model, this could be achieved in two non-mutually exclusive ways
\citep{quinones_Reinforcement_2019}: through low temporal discounting of
future effects and/or through perceiving a visitor client leaving as
psychological punishment (i.e.~as a negative reinforcer). Given that a
negative reinforcer is part of the ubiquitous learning mechanism termed
operant conditioning \citep{thorndike_Animal_1898, skinner_Behavior_1938},
this explanation can be
considered the simpler one. Furthermore, accounting for the future
effects of an action is considered an advanced cognitive ability
\citep{suddendorf_Evolution_2007}. Here, we fitted our model to the empirical
data on fish densities and cleaner fish performance in the market task
with the aim to determine the likely mechanism used by cleaners to
incorporate future consequences of current decisions.

\hypertarget{methods}{%
\subsection{Methods}\label{methods}}

\hypertarget{overview}{%
\subsubsection{Overview}\label{overview}}

We used the data from \citep{triki_Biological_2019} to fit the parameters of
the model analysed in \citep{quinones_Reinforcement_2019}. For the bayesian
fitting procedure we used Markov Chain Monte Carlo (MCMC) with a
Metropolis sampler to estimate the posterior distribution of the model
parameters. The data consist a of the number of times that a set of
cleaners chose to feed of a visitor client plate in the context of the
market task. This set of cleaners were collected in different locations
and in each one of those locations environmental variables were
measured. These environmental variables recorded from each location are
used as input for the computational model. Given these inputs, the
computational model produces a time series of choices in the market
task. In the model the preference for the two options changes as the
algorithm learns based on the reward that the two options provide. Thus,
we run the model for 10000 steps and use the last 3000 steps as the
preference developed in the model for visitors given the environmental
variables. We use this preference as the predicted probability in a
binomial distribution to calculate the likelihood that the number of
visitor choices obtained from each cleaner is generated by the model. We
sum over all the cleaners the log-likehoods and use the total for the
acceptance of new parameters in the MCMC.

\hypertarget{the-model}{%
\subsubsection{The model}\label{the-model}}

The model consist of a set individual-based simulations where
individuals face a series of choices between two options. Individuals in
the model represent cleaners, and the options they face simulate the
natural conditions of a cleaning market. Thus, individuals in the model
choose to prioritize different clients. These clients are classified as
resident and visitors. Residents are those clients that stay in the
cleaning queue when they are not given priority; while visitors leave
the queue (with a certain probability) after they are not given
priority. Individuals obtain a fixed reward from cleaning a client
regardless of the type. Every time individuals face and make a choice
they update the probability of making that same choice. The update is
based on the difference between the expectation of reward and the actual
reward the individual obtains (prediction error)
\citep{sutton_Reinforcement_2018, rescorla_Theory_1972}; and it is carried
in the direction that would lead to more reward being obtained, given
the new information. Further details of the model implementation can be
found in the paper \citep{quinones_Reinforcement_2019}. The model analysis
points to the relevance of two parameters. First, \(\gamma\) measures how
much individuals include future rewards in their decision updates. If
\(\gamma=0\), individuals only use the immediate reward obtained from a
cleaning interaction. As \(\gamma\) increases individuals include more the
reward obtained from the subsequent choices. That amounts to calculating
and using for decision making the future expected rewards in action.
Second, \(\eta\) measures how much individuals include in their reward the
fleeing behaviour of visitor clients as a negative component. Both of
these parameters allow individuals to use in their estimates the future
effects of their choices. The model analysis by Quiñones \emph{et al}
\citep{quinones_Reinforcement_2019} shows that without either of them, learning
individuals cannot develop a preference for the visitor client. The
analysis contrast the effect of these two parameters against variation
in the relative abundance of the two client types. The analysis shows
that under very high client (relative) abundance the presence of future
rewards (\(\gamma>0\)) no longer favours the preference for the ephemeral
clients, this effect does not occur with negative reward mediated by
client behaviour (\(\eta\)). Thus, the analysis suggests that future
reward as an explanation for cleaner ability to prefer visitors is
supported by Triki \emph{et al} \citep{triki_Biological_2019} data. However, a combination of
future and negative rewards, could still be a better explanation for the
data. In the current analysis, we will use the data from
\citep{triki_Biological_2019} to fit some of the model parameters, particularly
relevant we will fit \(\gamma\) and \(\eta\) to find out whether each or a
combination of these effects is a better explanation for the pattern
seen in the data. For other parameter values used for the model
simulations and the MCMC see Table \ref{tab:param}

\hypertarget{data-explanation}{%
\subsubsection{Data explanation}\label{data-explanation}}

\emph{Insert here brief description of the data and data collection}

In order to capture with the model the relationship between the
characteristics of the market (environment) and the cleaner's
performance we need to translate the absolute abundances of the data to
to a measure of relative abundances that captures client visitation patterns.
This is because in the model, relative abundances
of clients define not only the probability of residents and visitors,
but also how often the cleaning station is empty (e.g.~there are no
clients to be cleaned). The more cleaners there are, the more options
are available for the clients and hence the higher is the probability of
empty cleaning stations. How often the cleaning stations are empty in
turns influences the value that cleaners can expect from their choices.
The emptier the cleaning stations are, the more beneficial it is to
prioritize a client that will not stay. However, the frequency of empty
spots in the cleaning station does not only depend on cleaner abundances
but also on the frequency with which clients visit the
station. The client visitation frequency will likely vary among
different clients depending on parasite abundances and loads. These are
variables for which we do not have field estimates. Thus, we computed a
measure of relative cleaner abundance for each site relative to absolute
abundances and multiplied by a scaling constant that changes the range
of the variable. This scaling constant is meant to capture variation in the market
conditions driven partly by cleaner abundance. We fit the value of the
scaling constant as part of the
statistical inference. As for the client abundances for residents and
visitors we computed a relative measure with respect to the total client
abundance, and weighted that by the rescaled relative absence of
cleaners (1-relative cleaner abundance). Thus, all three measures of
relative abundance sum up to 1, and can be used in the model as a proxy
for the probability of having different options in the cleaning station.

\hypertarget{results}{%
\subsection{Results}\label{results}}

\begin{figure}

{\centering \includegraphics[width=0.95\linewidth]{manuscript_1.0_files/figure-latex/post-1}

}

\caption{Posterior distributions for parameter values $\gamma$, $\eta$ and the scaling constant. We show the kernel density estimates, below the mode (black point) and the 65\% (light blue) and 95\% (grey)  highest posterior density interval for the three paramteres. On the left (panels a, c and c) posterior distributions for a model with   all three parameters; while on the right (b and d) a model without negative reward. Panel f shows a measure of fit for both models, namely the distribution of pseudo-$R^2$ obtained from sampling the posterior distribution of parameter values.}\label{fig:post}
\end{figure}

Figure \ref{fig:post} panels a,c and e show the posterior distributions
of the parameter values of the full model, which includes both future
and negative reward. Posterior distributions show the probability
density (y axis) of the estimated parameter values (x axis). Notably,
the bulk of the posterior distribution for the parameter tuning negative
reward (\(\eta\) panel e) is around zero. Thus, the shape of the posterior
and its credible intervals suggest the absence of negative reward to be
well supported by the data. As for \(\gamma\), the confidence intervals
also includes zero, but the mode of the posterior is around 0.5. Thus,
we run the analysis setting \(\eta\) to zero. Figure \ref{fig:post} panels
b and d, show the posterior distributions of the two parameters
estimated. Not surprisingly, under a model without negative reward, the
distribution of \(\gamma\) shifts to higher values. Under the new model,
the estimates of \(\gamma\) do not include zero, the distribution is
centered around \(0.5\). To asses the fit of the two models we calculated
the \(pseudo-R^2\) \citep{mcfadden_Conditional_1974}, which is a standard
measure of fit for logistic regression. In that context, \(pseudo-R^2\)
uses the log-likelihood of the data given the model, relative to the
log-likelihood of the data given a model without covariates, as a
measure of fit. Given that our model is not a logistic regression, we
measure the \(pseudo-R^2\) in relation to the log-likelihood of a model
with parameters \(\gamma\) and \(\eta\) set to zero. This in practice
amounts to a model that has neutral preference between the two options.
In panel f of figure \ref{fig:post} we show the distribution of
\(pseudo-R^2\) calculated using samples from the posterior distributions
shown before, as well as from a model with negative reward but without
future reward. Note, \(pseudo-R^2\) can have negative values, that is when
the log-likelihood of the model is lower than that of a model that
triggers neutral preferences. Even though, the peak of the three
\(pseudo-R^2\) distributions is not very different, the model with only
future reward produces a distribution of \(pseudo-R^2\) where more values
are positive (to the right of black line in Fig. \ref{fig:post} f). This
shows that, accounting for variation in the parameter estimates, the
model with only future reward gives a better fit to the data, despite
having one parameter less.

\begin{figure}

{\centering \includegraphics[width=0.95\linewidth]{manuscript_1.0_files/figure-latex/pred-1}

}
